---
title: "30.Taxon_analysis"
author: "Christina Raw"
date: "11/5/2022"
output: 
  html_document:
    toc: true
---

Research supports that the response of biodiversity to agricultural systems can vary across taxa. In  this script I will test how different taxa respond to agricultural systems. 
 
Also, since I already know there is influential data, I am going to directly do the model with the robust package.

At the end of the script I make plots.

```{r, warning = FALSE, message = FALSE}

library(blme) # For Bayesian analysis
library(readxl) # For reading the excel data set
library(here) # For specifying file paths
library(knitr) # For making tables
library(tidyverse) # For data wrangling
library(sjPlot) # To make output tables
library(sjmisc) # To make output tables
library(sjlabelled) # To make output tables
library(ggplot2) # To plot

```

# 1.  Load data, take a look and prepare data frame

```{r, include = FALSE}

d <- read_excel(here("Datasets", "07.Excel_Dataset_to_model_LRR_LONG.xlsx"))

str(d)
d$Crop <- as.factor(d$Crop)
d$Treatment <- as.factor(d$Treatment)
d$Weight <- as.numeric(d$Weight)
d$ID <- as.factor(d$ID)

```

### 1.1. Data per kingdom

Kingdom is not really informative, so I will do it by phylum

```{r, echo = FALSE}
table(d$Kingdom)

```

### 1.2. Data per taxon

Curate phylum names and collapse categories

```{r}
d$Phylum[d$Phylum == "Insecta"] <- "Arthropoda"
d$Phylum[d$Phylum == "Azotobacter"] <- "Microbes"
d$Phylum[d$Phylum == "Microbial_community"] <- "Microbes"
d$Phylum[d$Phylum == "Microbiall_community"] <- "Microbes"
d$Phylum[d$Phylum == "Proteobacteria"] <- "Microbes"
d$Phylum[d$Phylum == "Actinomycetes"] <- "Microbes"
d$Phylum[d$Phylum == "Rhizosphere_actinobacteria"] <- "Rhizosphere_bacteria"
d$Phylum[d$Phylum == "Rhyzosphere_microbes"] <- "Rhizosphere_bacteria"
d$Phylum[d$Phylum == "Rhizosphere_Fungal_community"] <- "Rhizosphere_fungi"
d$Phylum[d$Phylum == "Rhizosphere_Fungal_population"] <- "Rhizosphere_fungi"
d$Phylum[d$Phylum == "Weeds"] <- "Weed"

Phylum <- as.data.frame(table(d$Phylum))
colnames(Phylum) <- c("Treatment", "N_Observations")
Phylum <- Phylum[order(Phylum$N_Observations, decreasing = TRUE),]

kable(Phylum, row.names = FALSE)


# Collapse organic and sustainable

d$Treatment[d$Treatment == "Organic"] <- "Sustainable"

# Include monoculture in conventional 

d$Treatment[d$Treatment == "Monoculture"] <- "Conventional"

# Make conventional the treatment reference level

d$Treatment <- relevel(d$Treatment, ref = "Conventional")


```

### 1.3. Subset phylum with over 10 obs

```{r}

phylum_obs <- table(d$Phylum) %>% 
  as.data.frame() %>% 
  subset(.$Freq >= 10)

d <- d[d$Phylum %in% phylum_obs$Var1, ]
```


# 2. Subset data by phylum

```{r, message = FALSE}

unique(d$Phylum)

group <- unique(d$Phylum)

group_list <- list() 

for (i in group){
data <- dplyr :: filter( d , Phylum == i)
group_list[[i]] <- data
}

list2env(lapply(group_list, as.data.frame), .GlobalEnv)

```

# 3. Within each phylum data frame, subset and model

Now that I have a separate data frame for each phylum, I will subset the agricultural systems that have more than 10 observations for that phylum. Then I will perform the linear models to see how functional groups respond to agricultural systems.

### 3.1. Arthropoda

```{r, echo = FALSE}
table(Arthropoda$Treatment)
```


```{r}

# Subset treatments that have more than 10 observations

treatment_obs <- table(Arthropoda$Treatment) %>% 
  as.data.frame() %>% 
  subset(.$Freq >= 10)

Arthropoda <- Arthropoda[Arthropoda$Treatment %in% treatment_obs$Var1, ]

# Model 

arthropoda_model <- blmer(LRR ~ Treatment + (1|ID) + (1|Crop), 
                     control = lmerControl(optimizer ="Nelder_Mead"), 
                     data = Arthropoda)


```

### 3.2. Microbes

This phylum does not have any agricultural treatment with more than 10 observations, so I will skip it. 

```{r, echo = FALSE}
table(Microbes$Treatment)
```


### 3.3. Mucoromycota

This phylum does not have enough data for analysis.

### 3.4. Rhizosphere bacteria

```{r}
# Subset treatments that have more than 10 observations

treatment_obs <- table(Rhizosphere_bacteria$Treatment) %>% 
  as.data.frame() %>% 
  subset(.$Freq >= 10)

Rhizosphere_bacteria <- Rhizosphere_bacteria[Rhizosphere_bacteria$Treatment %in% treatment_obs$Var1, ]
```

All observations come from one single paper, so not very useful but I will perform a lm nonetheless.

```{r}
# Model 

Rhizosphere_bacteria_model <- lm(LRR ~ Treatment, data = Rhizosphere_bacteria)

```

### 3.5. Rhizosphere fungi 

This phylum does not have enough data for analysis.

### 3.6. Vertebrates

```{r, error = TRUE}

# Subset treatments that have more than 10 observations

treatment_obs <- table(Vertebrates$Treatment) %>% 
  as.data.frame() %>% 
  subset(.$Freq >= 10)

Vertebrates <- Vertebrates[Vertebrates$Treatment %in% treatment_obs$Var1, ]

# Model 

Vertebrates_model <- blmer(LRR ~ Treatment + (1|ID) + (1|Crop), 
                     control = lmerControl(optimizer ="Nelder_Mead"), 
                     data = Vertebrates)

```

The error occurs because the vertebrate data set has data for uniquely oil palm, so I will remove crop random effect.

```{r}

Vertebrates_model <- blmer(LRR ~ Treatment + (1|ID), 
                     control = lmerControl(optimizer ="Nelder_Mead"), 
                     data = Vertebrates)
```

### 3.7. Weeds


```{r, error = FALSE}

# Subset treatments that have more than 10 observations

treatment_obs <- table(Weed$Treatment) %>% 
  as.data.frame() %>% 
  subset(.$Freq >= 10)

Weed <- Weed[Weed$Treatment %in% treatment_obs$Var1, ]

# Model 

Weed_model <- blmer(LRR ~ Treatment + (1|ID) + (1|Crop), 
                     control = lmerControl(optimizer ="Nelder_Mead"), 
                     data = Weed)

```


# 4. Model outputs

### 4.1. Blmer models

```{r, echo = FALSE}
tab_model(arthropoda_model, Vertebrates_model, Weed_model,
          dv.labels = c("Arthropoda", "Vertebrates", "Weeds"))
```

### 4.2. lm models

```{r, echo = FALSE}
tab_model(Rhizosphere_bacteria_model)

```

```{r, include = FALSE. eval = FASE}

# I am doing this to plot make a script where I plot everything

saveRDS(arthropoda_model,  here("Outputs/08.Results/arthropoda_model.rds"))
saveRDS(Vertebrates_model,  here("Outputs/08.Results/Vertebrates_model.rds"))
saveRDS(Weed_model,  here("Outputs/08.Results/Weed_model.rds"))
saveRDS(Rhizosphere_bacteria_model,  here("Outputs/08.Results/Rhizosphere_bacteria_model.rds"))

```

