---
title: "31.Functional_group_analysis"
author: "Christina Raw"
date: "24/5/2022"
output: 
  html_document:
    toc: true
---

In this script I am going to analyse how biodiversity responds to agricultural systems but taking a closer look at what is happening at the functional group level. Functional groups from [Brauman, *et. al*, (2020)](https://www.pnas.org/doi/10.1073/pnas.2010473117#t01)

Steps:

1. Subset the data by functional group
2. Within each functional group data frame, subset the treatment levels that have more than 10 observations.
3. Model

# 0. Load and prepare data frame

```{r, message = FALSE, warning = FALSE}

rm(list=ls())

library(readxl) # For reading the excel data set
library(here) # For specifying file paths
library(blme) # For Bayesian analysis
library(knitr) # For making tables
library(sjPlot) # For data wrangling
library(sjmisc) # For data wrangling
library(sjlabelled) # For data wrangling


d <- read_excel(here("Datasets", "07.Excel_Dataset_to_model_LRR_LONG.xlsx"))

d$Treatment <- as.factor(d$Treatment)
d$Functional_group <- as.factor(d$Functional_group)
d$ID <- as.factor(d$ID)
d$Crop <- as.factor(d$Crop)

# Collapse organic and sustainable

d$Treatment[d$Treatment == "Organic"] <- "Sustainable"

# Include monoculture in conventional 

d$Treatment[d$Treatment == "Monoculture"] <- "Conventional"

# Make conventional the treatment reference level

d$Treatment <- relevel(d$Treatment, ref = "Conventional")

```

# 1. Subset data by functional group

```{r, message = FALSE}

unique(d$Functional_group)

group <- unique(d$Functional_group)

group_list <- list() 

for (i in group){
data <- dplyr :: filter( d , Functional_group == i)
group_list[[i]] <- data
}

list2env(lapply(group_list, as.data.frame), .GlobalEnv)

```

# 2. Within each functional group data frame, subset and model

Now that I have a separate data frame for each functional group, I will subset the agricultural systems that have more than ten observations for that functional group. Then I will perform the linear models to see how functional groups respond to agricultural systems.

### 2.1. Air Climate Freshwater Soil and Extreme events

This functional group is basically represented by plants. Heads up, the following code might be annoying due to group's super long name.

```{r, echo = FALSE}
table(Air_Climate_Freshwater_Soil_ExtremeEvents$Treatment)
```


```{r}

# Subset treatments that have more than 10 observations

treatment_obs <- table(Air_Climate_Freshwater_Soil_ExtremeEvents$Treatment) %>% 
  as.data.frame() %>% 
  subset(.$Freq >= 10)

Air_Climate_Freshwater_Soil_ExtremeEvents <- Air_Climate_Freshwater_Soil_ExtremeEvents[Air_Climate_Freshwater_Soil_ExtremeEvents$Treatment %in% treatment_obs$Var1, ]

# Model 

plant_model <- blmer(LRR ~ Treatment + (1|ID) + (1|Crop), 
                     control = lmerControl(optimizer ="Nelder_Mead"), 
                     data = Air_Climate_Freshwater_Soil_ExtremeEvents)
summary(plant_model)

```

### 2.2. Borer

This functional group does not have any agricultural treatment with more than 10 observations, so I will skip it. 

```{r, echo = FALSE}

table(Borer$Treatment)

```


### 2.3. Natrual enemies

```{r, echo = FALSE}
table(Natural_enemy$Treatment)
```


```{r, error = TRUE}

# Subset treatments that have more than 10 observations

treatment_obs <- table(Natural_enemy$Treatment) %>% 
  as.data.frame() %>% 
  subset(.$Freq >= 10)

Natural_enemy <- Natural_enemy[Natural_enemy$Treatment %in% treatment_obs$Var1, ]

# Model 

enemy_model <- blmer(LRR ~ Treatment + (1|ID) + (1|Crop), 
                     control = lmerControl(optimizer ="Nelder_Mead"), 
                     data = Natural_enemy)
```

I found on [this link](https://community.rstudio.com/t/error-in-eval-f-x-downdated-vtv-is-not-positive-definite/73846) an explanation to this error: "This error means that the variance due to random effects is pretty small and close to zero". I went to the data and all the values for natural enemy LRR are = 0. I went back to the papers to double check whether it was a mistake. It is not. In the papers, they found no differences in natural enemy biodiversity between conventional and transgenic agricultural systems. 

So, for now, with this data set, I cannot study differences between agricultural systems for natural enemies.

### 2.4. Pest

```{r, echo = FALSE}

table(Pest$Treatment)
```


```{r}

# Subset treatments that have more than 10 observations

treatment_obs <- table(Pest$Treatment) %>% 
  as.data.frame() %>% 
  subset(.$Freq >= 10)

Pest <- Pest[Pest$Treatment %in% treatment_obs$Var1, ]

# Model 

pest_model <- blmer(LRR ~ Treatment + (1|ID) + (1|Crop), 
                    control = lmerControl(optimizer ="Nelder_Mead"), 
                    data = Pest)

summary(pest_model)
```


### 2.5. Pollination and seed dispersal

This functional group does not have any agricultural treatment with more than 10 observations, so I will skip it. 

```{r, echo= FALSE}
table(Pollination_seeds$Treatment)
```


### 2.6. Predators 

```{r, echo = FALSE}
table(Predators$Treatment)
```


```{r, error = TRUE}

# Subset treatments that have more than 10 observations

treatment_obs <- table(Predators$Treatment) %>% 
  as.data.frame() %>% 
  subset(.$Freq >= 10)

Predators <- Predators[Predators$Treatment %in% treatment_obs$Var1, ]

# Model

predator_model <- blmer(LRR ~ Treatment + (1|ID) + (1|Crop), 
                        control = lmerControl(optimizer ="Nelder_Mead"), 
                        data = Predators)

```

Here the same thing happens as with natural enemies, all the LRR observations are = 0. So, for now, with this data set, I cannot study differences between agricultural systems for predators.


### 2.7. Soil

```{r, echo = FALSE}
table(Soil$Treatment)
```


```{r}

# Subset treatments that have more than 10 observations

treatment_obs <- table(Soil$Treatment) %>% 
  as.data.frame() %>% 
  subset(.$Freq >= 10)

Soil <- Soil[Soil$Treatment %in% treatment_obs$Var1, ]

# Model 

soil_model <- blmer(LRR ~ Treatment + (1|ID) + (1|Crop), 
                        control = lmerControl(optimizer ="Nelder_Mead"), 
                        data = Soil)
summary(soil_model)

# Save model to make summary tables and plots

saveRDS(soil_model,  here("Outputs/08.Results/soil_model.rds"))


```


### 2.8. Soil and pollination

Not enough observations

# 3. Model outputs

```{r, echo = FALSE}
tab_model(pest_model, plant_model, soil_model,
          dv.labels = c("Pests", "AirClimateFreshwaterSoilExtremeEvents", "Creation and protection of Soil"))

```

