---
title: "Crops and livestock broad scoping phase"
author: "Christina Raw"
date: "16/11/2021"
output: html_document
---

Overall aim: To map out the state of knowledge of the impacts of widespread agricultural land-use systems on local biodiversity. This will identify any important gaps and will also provide a scaffold on which we can build.

Step 1: From FAO, get the most recent data available on area used globally to grow different crops and to farm different livestock species; and also the global value of these crops and species.

  1) Broad scope: download information from FAOSTAT to see which are the global top 15 crops and the global top 5 livestock species.
  
  2) Detailed scope: within the top crops and livestock species, search research syntheses on the main farming systems. 


# Crops broad scope

I downloaded the data sets from [FAOSTAT](https://www.fao.org/faostat/en/#data). Specifically, I downloaded:

  1. Production
  
      1.1. Crops and livestock products. [Metadata](https://www.fao.org/faostat/en/#data/QCL/metadata)
      
      1.2. Value of agricultural production. [Metadata](https://www.fao.org/faostat/en/#data/QV/metadata)
  
### 1.1. Crops and livestock products.

This data set contains information about the global harvested area, yield and production of each crop in 2019. 
To obtain the top 15 crops farmed globally, first I subset the data set by area harvested. Then, I sorted by value in descending order. Finally, I subset the first 15 rows, which are the top 15 crops farmed globally.


```{r, include = FALSE}

# Load data set 

crops_primary_total <- read.csv("Data/00.Raw_Data/00.Scoping/Crops_and_livestock_products/Crops/FAOSTATcrops_primary_total_2019.csv")

colnames(crops_primary_total)

# Subset by area harvested, yield and production

unique(crops_primary_total$Element)

crop_primary_area <- subset(crops_primary_total, crops_primary_total$Element == "Area harvested")
crop_primary_yield <- subset(crops_primary_total, crops_primary_total$Element == "Yield")
crops_primary_production <- subset(crops_primary_total, crops_primary_total$Element == "Production")

# Sort by value in descending order and subset first 15 rows, which are the top 15 crops

crop_primary_area <- crop_primary_area[order(-crop_primary_area$Value),]
crop_primary_yield <- crop_primary_yield[order(-crop_primary_yield$Value),]
crops_primary_production <- crops_primary_production[order(-crops_primary_production$Value),]

area <- crop_primary_area[1:15,]
yield <- crop_primary_yield[1:15,]
production <- crops_primary_production[1:15,]


```

```{r, echo = FALSE}

# Create a table of the top 15 crops farmed globally by area

top_15_crops <- data.frame(area$Item, area$Value)
colnames(top_15_crops) <- c("Crop","Hectares")
save(top_15_crops, file= "Outputs/Top_15_crops_global_area.Rda")

library(knitr)

kable(top_15_crops, align = "l", caption = "Top 15 crops farmed globally by area", row.names = TRUE)

```

    
### 1.2. Value of agricultural production.
 
This data set contains information about the global value of gross production of farmed crops. I selected the value of gross production of the top 15 crops farmed globally. 


```{r, include = FALSE}
 
value_total <- read.csv("Data/00.Raw_Data/00.Scoping/Value_agricultural_production/FAOSTAT_value_agricultural_products_total_2019.csv")
 
# Names of the top 15 crops farmed globally

top_15_crops$Crop

# Subset the gross production value of the top 15 crops farmed globally

library(dplyr)

top_15_value <- subset(value_total, value_total$Item == "Wheat" |value_total$Item == "Maize" |value_total$Item == "Rice, paddy" | value_total$Item == "Soybeans"| value_total$Item == "Barley" |value_total$Item == "Sorghum"|value_total$Item == "Seed cotton"|value_total$Item == "Rapeseed"|value_total$Item == "Beans, dry"|value_total$Item == "Millet"|value_total$Item == "Groundnuts, with shell"|value_total$Item == "Oil palm fruit"|value_total$Item == "Cassava"|value_total$Item == "Sunflower seed"|value_total$Item == "Sugar cane" )


# Create a table of the top 15 crops farmed globally by area and their gross production value

    # Select columns of interest

top_15_value <- data.frame(top_15_value$Item, top_15_value$Value)
colnames(top_15_value) <- c("Crop","1000 Int. $")

    # Join both data frames 

top_15_crop_value <- inner_join(top_15_crops, top_15_value)


# Add commodity category. I obtained the commodities from FAO

Commodity <- c(rep("Cereals & grains", 6), rep("Oils & fats", 2), "Legumes", "Cereals & grains", rep("Oils & fats", 2), "Roots & tubers", "Oils & fats", "Sugar")

top_15_crop_value <- cbind(top_15_crop_value, Commodity)

write.csv(top_15_crop_value, "Outputs/00.Crops/Top_15_crop_value.csv")
write.csv(top_15_crop_value, "Data/01.Processed_Data/00.Crops/Top_15_crop_value.csv")

```

```{r, echo = FALSE}
 
# Make table 

top_15_crop_value <- read.csv("Outputs/00.Crops/Top_15_crop_value.csv", row.names = 1)

library(knitr)

kable(top_15_crop_value, caption = "Top 15 crops farmed globally by area and value of gross production", row.names = TRUE)
```


# Cattle broad scope

All the information I have found about livestock either talks about number of heads or livestock density (individuals per area).

### Number of heads: 
  
  - [Sandhage-Hofmann, A. 2016](https://www.sciencedirect.com/topics/agricultural-and-biological-sciences/livestock-species)
  
```{r, echo = FALSE}

species <- data.frame(Species = c("Cattle", "Sheep", "Goat", "Horses", "Camels", "Donkeys"), Value= c(1.4, 1.1, 0.8, 56.4, 22, 44), Unit = c("billion", "billion", "billion", "million", "million", "million"))

write.csv(species, "Outputs/01.Livestock/Sandhage-Hofmann.csv")

library(knitr)

kable(species, caption = "Sandhage-Hofmann, A. (2016) livestock data", align = "l", row.names = FALSE)
```

  - [FAOSTAT](https://www.fao.org/faostat/en/#data). I used two data sets. One data set contains information about global livestock production in 2019 and. The remaining data set contains information about the global value of agricultural production in 2019 from which I selected the value of gross production of the livestock produced in 2019.  
      
```{r, include = FALSE}

library(tidyr)
library(dplyr)

# Load data set of livestock production and select columns of interest

livestock <- read.csv("Data/00.Raw_Data/00.Scoping/Crops_and_livestock_products/Livestock/FAOSTAT_livestock_primary_total_2019.csv") 

livestock <- subset(livestock, livestock$Unit =="No"| livestock$Unit == "1000 Head" | livestock$Unit  == "Head")  

livestock <- select(livestock, 8, 11, 12)

# Load data set of value of agricultural production

livestock_value <- read.csv("Data/00.Raw_Data/00.Scoping/Value_agricultural_production/FAOSTAT_value_agricultural_products_total_2019.csv")

livestock$Cost <- ""

# Add the production value from to the livestock species of the livestock production data set

for (i in (1:186)){
  
  for (j in (1:43)){
    if (livestock_value$Item[i] == livestock$Item[j]){
      livestock$Cost[j] <- livestock_value$Value[i]
    }}}

colnames(livestock)[4]<-"1000 Int. $"


# Sort by value in descending order

livestock <- livestock [order(-livestock$Value),]

write.csv(livestock, "Outputs/01.Livestock/Livestock_value.csv")
write.csv(livestock, "Data/01.Processed_Data/01.Livestock/Livestock_value.csv")

```

```{r, echo = FALSE}
library(knitr)

kable(livestock, caption = "FAOSTAT livestock data", row.names = FALSE)

```

### Density:
  
  - Gridded Livestock of the World [GLW3](https://dataverse.harvard.edu/dataverse/glw_3). 

Maps global livestock  density distribution in raster layers, as number of individuals per pixel.  I tried to vectorise the raster layer and calculate the total area occupied by livestock production but this is pointless because there the area is basically the whole world, as there is a raster level of 0-1 individuals/pixel. Then I thought of transforming to presence-absence raster and then vectorise to calculate the area. Problem: where is the livestock density cutoff? It's arbitrary, not a good approach. 
  
  - Timothy, R., et al. 2014 [PLOS](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0096084#pone-0096084-g002):
  
"The highest cattle densities (Figure 2a) are found in India, in the East African highlands (particularly in Ethiopia), in Northern Europe and in South America. Desert areas and the tropical rain forests of Amazonia and of the Congo Basin have practically no cattle. The highest concentrations of pigs are found in China and in other Eastern Pacific countries (Figure 2b). Pigs are also densely distributed in European countries while only a few countries in Africa (e.g. Uganda, Burkina Faso, Ghana, Nigeria and Togo) have significant densities. Relatively high concentrations are also found in Central America and in Brazil. The distribution of chickens (Figure 2c) closely follows that of the human population. The highest concentrations are found in eastern China, in Pakistan and India, and in western Europe. In Africa, the countries facing the Gulf of Guinea and Madagascar also have high chicken densities. The densely populated east coast of the United States also shows high numbers whilst chickens are only sparsely distributed in the central and western states. The heavily populated areas of southern Brazil also show high concentrations of chickens. The distribution of ducks (Figure 2d), for those regions for which sub-national statistics were available, adds information to previous national and regional duck mapping efforts [18], [19]. Ducks are far less common than chickens worldwide although high densities are to be found in South-East Asia and China where duck production is often integrated with rice cropping and fish farming [56]–[57]. Though to a lesser extent, duck densities are also quite high in a few European countries (e.g. France). file S4 provides a summary of the sub-national statistics used for the modelling and file S5 provides detailed metadata for the sub-national statistics used to develop the livestock distribution models. file S6 provides two graphic summaries of data availability for the modelled species (a) the average spatial resolution of the training data and (b) administrative level of the training data."
      
     
# Conclusions

  - I will study the top 15 crops mentioned in the crop broad scope

  - I think I will study the main livestock species mentioned by FAO and GL3: sheep, pig, horse, goat, duck, cattle, chicken and buffalo. I believe that if there is enough information to map their global density distribution, there will be enough information to study their systems of production. 

  - I have to talk about cattle in terms of number of heads or density and not in terms of area.


      
      
      


