---
title: "Data digest"
author: "Christina Raw"
date: "9/2/2022"
output: html_document
---

```{r, include = FALSE}
library(here)

```


In  this html I am going to digest the crop data I have extracted from syntheses on agricultural production systems
on biodiversitt. The aim is to provide an easy overview of the data set. I am going to classify the data  according to VividEconomics land-use classes.  

The **data set**  consists of the data I have extracted from the included papers, data on the impact
of different production systems of the selected crops on biodiversity. The data are quantitative and qualitative
metrics of biodiversity, such as species richness or species abundance under different agricultural production systems. 

VividEconomics has the following cropland land use classes:
  
  - Annual crops: rice, maize, tropical cereals, temperate cereals, cotton seed, groundnuts, sunflower, other oil crops (including rape seed), fruits, vegetables and nuts.
  
  - Perennial crops: sugar beet, sugar cane, short rotation grasses (bioenergy), short rotation trees (bioenergy), potato, oil palm and tropical roots.
    
  - N-fixing crops: pulses and soy.  
   
According to VividEconomics LU classes, I am going to count:
   
  a. Number of annual, perennial and N-fixing crops
  b. Within each group, number of records per crop
   
I am going to do this for each of the two types of data in the data set:
  
  a. Quantitative data
  b. Qualitative data 
  
## a. Quantitative data
   
### Number of annual, perennial and N-fixing crops 


```{r, include = FALSE}

# How many records per aggregated land use classes

# 1. Subset by LU classes according to Vivid Economics: annual, perennial and N-Fixing crops.
# 2. Add up the records for aggregated land use classes
# 3. Make table for aggregated land use-classes


library(dplyr)
library(here)

d <- read.csv(here("Data/01.Processed_Data/03.Curated_spreadsheet", "Csv_Crops_Quantitative_spreadsheet.csv"), row.names = 1) # Load data


# 1. Subset by LU classes according to Vivid Economics: annual, perennial and N-Fixing crops 
# 2. Add up the records for aggregated land use classes

names(d)
str(d)
unique(d$Crop)

     # Annual crops

annual_crops <- subset(d, d$Crop == "Rice" | d$Crop == "Bt_Rice" |
                          d$Crop == "Maize" | d$Crop == "Bt_Maize" |
                          d$Crop == "Barley" | d$Crop == "Straw_cereals" |
                          d$Crop == "Small_grain" | d$Crop == "Sorghum" |
                          d$Crop == "Wheat" | d$Crop == "Cotton" |
                          d$Crop == "Bt_Cotton" | d$Crop == "GM_Cotton" |
                          d$Crop == "Sunflower" | d$Crop == "Bt_Sunflower" |
                          d$Crop == "Rape_seed" | d$Crop == "Bt_Rape" |
                          d$Crop == "GM_Rape" | d$Crop == "Bt_Corn" |
                          d$Crop == "Bt_Eggplant" | d$Crop == "Bt_Broccoli" |
                          d$Crop == "GM_Tomato" | d$Crop == "Gm_Strawberry")

length(annual_crops$Crop) # 349 records for annual crops

length(unique(annual_crops$Paper_ID)) # 17 papers

        
        # Perennial crops

perennial_crops <- subset(d, d$Crop == "Sugarbeet" | d$Crop == "Sugarcane" |
                             d$Crop == "Potato" | d$Crop == "Bt_Potato"| d$Crop == "GM_Potato" |
                             d$Crop == "Oil_palm" | d$Crop ==  "Cassava")
                              
length(perennial_crops$Crop) # 214 records for perennial crops
length(unique(perennial_crops$Paper_ID)) # 18 papers


        # N-Fixing crops
                              
N_fixing_crops <- subset(d, d$Crop == "Soybean" | d$Crop == "Maize_Legume" |
                             d$Crop == "Legume")

length(N_fixing_crops$Crop) # 25 records for N-Fixing crops
length(unique(N_fixing_crops$Paper_ID)) # 5 papers

```

```{r, echo = FALSE }

# 3. Make table for aggregated land use-classes

LU <- c("Annual crops", "Perennial crops", "N-fixing crops")
sample_size <- c(349, 214, 25)
papers <- c(17, 18, 5)

aggregated <- data.frame(LU, sample_size, papers)

colnames(aggregated) <- c("Cropland land-use", "Sample size", "Number of papers")

library(knitr)

kable(aggregated, caption = "Data for VividEconomics QUANTITATIVE aggregated cropland land-uses", row.names = TRUE)

```


### Within each group, number of records per crop

```{r, include = FALSE}

# Steps: 

# 1. Add up the records dissagregated land-use classes 
# 2. Add the number of papers per crop as a column in the data frame 
# 3. Make a table with all the information


# 1. Add up the records dissagregated land-use classes 


        # Table of crops per aggregated land-use

disagregated_annual <- as.data.frame(table(annual_crops$Crop))
disagregated_perennial <- as.data.frame(table(perennial_crops$Crop))
disagregated_N_fixing <- as.data.frame(table(N_fixing_crops$Crop))
       
        # Add column of aggregated category to each of the disaggregated data 
        # frame, i.e., add column that says "annual crops" 
        # to the data frame of aggregated annual crops. And change 

disagregated_annual <- cbind(disagregated_annual, Aggregated_LU = "Annual crops")  
disagregated_perennial <- cbind(disagregated_perennial, Aggregated_LU = "Perennial crops") 
disagregated_N_fixing <- cbind(disagregated_N_fixing, Aggregated_LU = "N-fixing crops") 

        # Now, rbind to make a nice table

crop_table <- rbind(disagregated_annual, disagregated_perennial, disagregated_N_fixing)


# 2. Add the number of papers per crop as a column in the data frame 
#    Now I want to count how many papers per crop I there are to, ultimately, make a cool table with 
#    number of papers, number of data per commodity, number of records per crop

library(dplyr)

papers <- select(d, Paper_ID, Crop) # I'm going to subset columns of interest to count
                                    # papers there are per crop

unique_papers <- unique(papers) # remove duplicate rows so I have unique paper names for the crop 

n_papers <- as.data.frame(table(unique_papers$Crop)) # Table so I can count the number oc papers per crop



colnames(n_papers)[2] <- "Number of papers" # First. I am gong to change the name of the column to avoid confusions

order <- crop_table$Var1                             # Order the rows to match the crop_table 
n_papers <- n_papers[match(order, n_papers$Var1),]   # data frame row order

crop_table <- cbind(crop_table, n_papers)            # Add the number of papers per crop as a column in the df

colnames(crop_table)[1] <- "Crop"           # Change column names, reorder them, drop unwanted
colnames(crop_table)[2] <- "Sample_size"
crop_table <- crop_table[, -4]

colnames(crop_table)
crop_table <- crop_table[, c(3, 1, 2, 4)]

crop_table <- arrange(crop_table, desc(crop_table$Sample_size)) # Sort by sample size descending order


```


```{r, echo = FALSE }

# 3. Make table

kable(crop_table, caption = "Data for VividEconomics QUANTITATIVE disaggregated cropland land-uses", row.names = TRUE)
```


## b. Qualitative data

```{r, include = FALSE}

# How many records per aggregated land use classes

# 1. Subset by LU classes according to Vivid Economics: annual, perennial and N-Fixing crops.
# 2. Add up the records for aggregated land use classes
# 3. Make table for aggregated land use-classes


library(dplyr)
library(here)

d <- read.csv(here("Data/01.Processed_Data/03.Curated_spreadsheet",  "Csv_Crops_Qualitative_spreadsheet.csv"), row.names = 1) # Load data


# 1. Subset by LU classes according to Vivid Economics: annual, perennial and N-Fixing crops 
# 2. Add up the records for aggregated land use classes

names(d)
str(d)
unique(d$Crop)

     # Annual crops

annual_crops <- subset(d, d$Crop == "Bt_Rice" | d$Crop == "Rice" |
                          d$Crop == "Bt_Maize" | d$Crop == "Maize" |
                          d$Crop == "Gr_Maize" | d$Crop == "Wheat" |
                          d$Crop == "Barley" | d$Crop == "Sorghum" |
                          d$Crop == "Bt_Cotton" | d$Crop == "Cotton" |
                          d$Crop == "Corn" | d$Crop == "Bt_Corn" |
                          d$Crop == "Rape_seed" | d$Crop == "Gr_Rape_seed" |
                          d$Crop == "GM_Rape_seed" | d$Crop == "Bt_Tobacco")

length(annual_crops$Crop) # 211 records for annual crops
length(unique(annual_crops$Paper_ID)) # 14 papers

        
        # Perennial crops

perennial_crops <- subset(d, d$Crop == "Sugarbeet" | d$Crop == "Sugarcane" |
                             d$Crop == "Gr_Sugarbeet" | d$Crop == ""| d$Crop == "GM_Sugarcane" |
                             d$Crop == "GM_Potato" | d$Crop ==  "Oil_palm")

length(perennial_crops$Crop) # 189 records for perennial crops
length(unique(perennial_crops$Paper_ID)) # 15 papers



        # N-Fixing crops
                              
N_fixing_crops <- subset(d, d$Crop == "Soybean" | d$Crop == "Bt_Soybean" |
                             d$Crop == "Faba_bean")

length(N_fixing_crops$Crop) # 40 records for N-Fixing crops
length(unique(N_fixing_crops$Paper_ID)) # 4 papers

```

```{r, echo = FALSE }

# 3. Make table for aggregated land use-classes

LU <- c("Annual crops", "Perennial crops", "N-fixing crops")
sample_size <- c(211, 189, 40)
papers <- c(14, 15, 4)

aggregated <- data.frame(LU, sample_size, papers)

colnames(aggregated) <- c("Cropland land-use", "Sample size", "Number of papers")

library(knitr)

kable(aggregated, caption = "Data for VividEconomics QUALITATIVE aggregated cropland land-uses", row.names = TRUE)

```


### Within each group, number of records per crop

```{r, include = FALSE}

# Steps: 

# 1. Add up the records dissagregated land-use classes 
# 2. Add the number of papers per crop as a column in the data frame 
# 3. Make a table with all the information


# 1. Add up the records dissagregated land-use classes 


        # Table of crops per aggregated land-use

disagregated_annual <- as.data.frame(table(annual_crops$Crop))
disagregated_perennial <- as.data.frame(table(perennial_crops$Crop))
disagregated_N_fixing <- as.data.frame(table(N_fixing_crops$Crop))

       
        # Add column of aggregated category to each of the disaggregated data 
        # frame, i.e., add column that says "annual crops" 
        # to the data frame of aggregated annual crops. And change 

disagregated_annual <- cbind(disagregated_annual, Aggregated_LU = "Annual crops")  
disagregated_perennial <- cbind(disagregated_perennial, Aggregated_LU = "Perennial crops") 
disagregated_N_fixing <- cbind(disagregated_N_fixing, Aggregated_LU = "N-fixing crops") 


        # Now, rbind to make a nice table

crop_table <- rbind(disagregated_annual, disagregated_perennial, disagregated_N_fixing)


# 2. Add the number of papers per crop as a column in the data frame 
#    Now I want to count how many papers per crop I there are to, ultimately, make a cool table with 
#    number of papers, number of data per commodity, number of records per crop

library(dplyr)

papers <- select(d, Paper_ID, Crop) # I'm going to subset columns of interest to count
                                    # papers there are per crop

unique_papers <- unique(papers) # remove duplicate rows so I have unique paper names for the crop 

n_papers <- as.data.frame(table(unique_papers$Crop)) # Table so I can count the number oc papers per crop



colnames(n_papers)[2] <- "Number of papers" # First. I am gong to change the name of the column to avoid confusions

order <- crop_table$Var1                             # Order the rows to match the crop_table 
n_papers <- n_papers[match(order, n_papers$Var1),]   # data frame row order

crop_table <- cbind(crop_table, n_papers)            # Add the number of papers per crop as a column in the df

colnames(crop_table)[1] <- "Crop"           # Change column names, reorder them, drop unwanted
colnames(crop_table)[2] <- "Sample_size"
crop_table <- crop_table[, -4]

colnames(crop_table)
crop_table <- crop_table[, c(3, 1, 2, 4)]

crop_table <- arrange(crop_table, desc(crop_table$Sample_size)) # Sort by sample size descending order

```


```{r, echo = FALSE }

# 3. Make table

kable(crop_table, caption = "Data for VividEconomics QUALITATIVE disaggregated cropland land-uses", row.names = TRUE)
```














