---
title: "CropQuantitativeDataDigest"
author: "Christina Raw"
date: "7/3/2022"
output: html_document
---


## Number of observations per magpie class and, within class, number of observations per production system

There is enough data to compare production systems for cereals, oil crops, and sugar.
Maybe it could also be done for other annual crops.

```{r, message = FALSE,  echo = FALSE}

rm(list=ls())
library(dplyr)
library(writexl)

library(here)
d <- read.csv(here("Data/01.Processed_Data/04.Magpie_classes",  "Csv_Magpie_Crops_Quantitative_spreadsheet.csv"), row.names = 1) # Load data

magpie_data <- as.data.frame(table(d$magpie_class))

colnames(magpie_data) <- c("Magpie class", "N observations" )

library(knitr)

# print the results

kable(magpie_data, row.names = TRUE)

```



```{r, message = FALSE,  echo = FALSE}

library(tidyr)

d$Crop[d$Crop == "Bt_Corn"]<- "Corn"
d$Crop[d$Crop == "Bt_Potato"]<- "Potato"
d$Crop[d$Crop == "Bt_Cotton"]<- "Cotton"
d$Crop[d$Crop == "Bt_Eggplant"]<- "Eggplant"
d$Crop[d$Crop == "Bt_Rice"]<- "Rice"
d$Crop[d$Crop == "Bt_Maize"]<- "Maize"
d$Crop[d$Crop == "Bt_Sunflower"]<- "Sunflower"
d$Crop[d$Crop == "Bt_Rape"]<- "Rape_seed"
d$Crop[d$Crop == "GM_Rape"]<- "Rape_seed"
d$Crop[d$Crop == "Bt_Broccoli"]<- "Broccoli"
d$Crop[d$Crop == "GM_Potato"]<- "Potato"
d$Crop[d$Crop == "GM_Cotton"]<- "Cotton"
d$Crop[d$Crop == "GM_Tomato"]<- "Tomato"
d$Crop[d$Crop == "Gm_Strawberry"]<- "Strawberry"


magpie_reshape <- select(d, magpie_class, agricultural_system)
magpie_reshape <- as.data.frame(table(magpie_reshape))

magpie_reshape <- magpie_reshape %>% group_by(magpie_class) %>% pivot_wider(names_from = agricultural_system,
                                          values_from = Freq)
kable(magpie_reshape, row.names = TRUE)

```


## For each crop, number of observations and papers 

```{r, message = FALSE,  echo = FALSE}

# 1. Add magpie classes column

magpie <- read.csv2(here("Data/01.Processed_Data/04.Magpie_classes", "CSV_Magpie_classess.csv"))

crop_data <- as.data.frame(table(d$Crop))

  # Add column with magpie classes
colnames(crop_data)[1] <- "Crop"
colnames(crop_data)[2] <- "N"

colnames(magpie)[3] <- "My_crops"  # Change column name that bothered me

#Input magpie classes into my d data set
magpie <- na.omit(magpie)  # First I remove rows with NA in the magpie data set
                           # for two reasons: 
                           # 1. NA rows means I don't have data for that crop
                           # 2. NA's will interrupt the following for loop

crop_data$Magpie_class <- ""  # Create column where I will store the info

for (i in (1:22)){
  #browser()
  for (j in (1:31)){
    if (crop_data$Crop[i] == magpie$My_crops[j]){
      crop_data$Magpie_class[i] <- magpie$Magpie_classess[j]
    }}}

# Correct barley
magpie[5,3] <- "Barley"

# Rerun loop 

for (i in (1:22)){
  #browser()
  for (j in (1:31)){
    if (crop_data$Crop[i] == magpie$My_crops[j]){
      crop_data$Magpie_class[i] <- magpie$Magpie_classess[j]
    }}}


crop_data <- relocate(crop_data, Magpie_class, .before = Crop) # Reorder columns

# 2. Count papers per crop

papers <- select(d, Crop, Paper_ID)

data_count_1 <- aggregate(data = papers,                  # Count the number of 
                          Paper_ID ~ Crop,                # per crop
                          function(x) length(unique(x)))

crop_Nstudies_Npapers <- merge(crop_data,data_count_1, by="Crop")  # Make table

    #Change colnames and oder in descending number of observations

colnames(crop_Nstudies_Npapers)[c(3,4)] <- c("N", "Number of papers" )

crop_Nstudies_Npapers <- crop_Nstudies_Npapers[order(-crop_Nstudies_Npapers$N),]


# 3. Make table

kable(crop_Nstudies_Npapers, row.names = TRUE)

```

## For each crop, number of observations per agricultural system

If I get all the data to be of the same type, I could compare:

  - Cotton: mixed and transgenic
  - Maize: conservation, mixed, transgenic
  - Potato (potentially): conventional, organic, transgenic
  - Rapeseed: IPM and transgenic
  - Sugarcane: conservation and conventional
  - Wheat: conventional and organic

```{r,  message = FALSE,  echo = FALSE}

library(tidyr)

# Magpie reshape 

magpie_reshape <- select(d, Crop, agricultural_system)
magpie_reshape <- as.data.frame(table(magpie_reshape))

magpie_reshape <- magpie_reshape %>% group_by(Crop) %>% pivot_wider(names_from = agricultural_system,
                                          values_from = Freq)
kable(magpie_reshape, row.names = TRUE)

```

## Number of observations per effect size type

```{r, echo = FALSE}

d$Effect_size[d$Effect_size == "Hedges_d"]<- "Hedges’ d"
es <- as.data.frame(table(d$Effect_size))

colnames(es) <- c("Effec size", "N of observations")
es <- es[order(-es$`N of observations`),]

kable(es, row.names = TRUE)

efs <- aggregate(data = d,                        # Count the number of papers
               Effect_size ~ Crop,        # per agricultural system
               function(x) length(unique(x)))

```

## For each crop, number of effect size observations

```{r, message = FALSE, echo = FALSE}

library(tidyr)

d$Effect_size[d$Effect_size == "Hedges’ d"]<- "Hedges' d"
d$Effect_size[d$Effect_size == "Hedges_d"]<- "Hedges' d"

reshape <- select(d, Crop, Effect_size)
reshape <- as.data.frame(table(reshape))

reshape <- reshape %>% group_by(Crop) %>% pivot_wider(names_from = Effect_size,
                                          values_from = Freq)

kable(reshape, row.names = TRUE)


```


```{r,eval = FALSE, echo = FALSE}

# Save data frames

write_xlsx(magpie_data, "Outputs/06.Magpie_classes/DataDigest/Excel_DataDigest_MagpieClasses.xlsx")
write.csv(magpie_data, "Outputs/06.Magpie_classes/DataDigest/CSV_DataDigest_MagpieClasses.csv")

write_xlsx(crop_data, "Outputs/06.Magpie_classes/DataDigest/Excel_DataDigest_CropData.xlsx")
write.csv(crop_data, "Outputs/06.Magpie_classes/DataDigest/CSV_DataDigest_CropData.csv")

write_xlsx(crop_Nstudies_Npapers, "Outputs/06.Magpie_classes/DataDigest/Crops_Nstudies_Npapers.xlsx")
```


## For each prodcution system, number of observations and papers  

Potentially useful to do an analysis per agricultural system. It seems as there could be 
enough data for all production systems except unclassified and IPM.

```{r, include = FALSE}

systems <- aggregate(data = d,                        # Count the number of papers
               Paper_ID ~ agricultural_system,        # per agricultural system
               function(x) length(unique(x)))                                     
               

systems_obs <- as.data.frame(table(d$agricultural_system))
colnames(systems_obs) <- c("agricultural_system", "N of observations")

systems <- merge( systems_obs, systems, by = "agricultural_system")
systems <- systems[order(-systems$`N of observations`),]

colnames(systems) <- c("agricultural_system", "N of observations", "N of papers")

```



```{r, echo = FALSE}


kable(systems, row.names = TRUE)

```




# Conclusion

There is enough data to compare production systems for cereals, oil crops, and sugar. Maybe it could also be done for other annual crops.

To perform a meta analysis, I would need to extract the acual effect size and variance for 
all the percentage observations. I will have a read on other options.

If I get all the data to be of the same type, I could compare:

Cotton: mixed and transgenic
Maize: conservation, mixed, transgenic
Potato (potentially): conventional, organic, transgenic
Rapeseed: IPM and transgenic
Sugarcane: conservation and conventional
Wheat: conventional and organic

Potentially useful to do an analysis per agricultural system. It seems as there could be 
enough data for all production systems except unclassified and IPM,
