---
title: "14.VividEconomics_data_digest"
author: "Christina Raw"
date: "8/3/2022"
output: html_document
---

## Number of observations per magpie class and, within class, number of observations per production system

```{r, include = FALSE}

library(dplyr)
library(here)
d <- read.csv(here("Data/01.Processed_Data/04.Magpie_classes",  "Csv_Magpie_Crops_Quantitative_spreadsheet.csv"), row.names = 1) # Load data

d$Crop[d$Crop == "Bt_Corn"]<- "Corn"
d$Crop[d$Crop == "Bt_Potato"]<- "Potato"
d$Crop[d$Crop == "Bt_Cotton"]<- "Cotton"
d$Crop[d$Crop == "Bt_Eggplant"]<- "Eggplant"
d$Crop[d$Crop == "Bt_Rice"]<- "Rice"
d$Crop[d$Crop == "Bt_Maize"]<- "Maize"
d$Crop[d$Crop == "Bt_Sunflower"]<- "Sunflower"
d$Crop[d$Crop == "Bt_Rape"]<- "Rape_seed"
d$Crop[d$Crop == "GM_Rape"]<- "Rape_seed"
d$Crop[d$Crop == "Bt_Broccoli"]<- "Broccoli"
d$Crop[d$Crop == "GM_Potato"]<- "Potato"
d$Crop[d$Crop == "GM_Cotton"]<- "Cotton"
d$Crop[d$Crop == "GM_Tomato"]<- "Tomato"
d$Crop[d$Crop == "Gm_Strawberry"]<- "Strawberry"

d <- subset(d, d$Crop == "Maize"|
            d$Crop == "Cotton"|
            d$Crop == "Oil_palm"|
            d$Crop == "Soybean"|
            d$Crop == "Sugarcane"|
            d$Crop == "Strawberry"|
            d$Crop == "Broccoli"|
            d$Crop == "Tomato"|
            d$Crop == "Potato"|
            d$Crop == "Eggplant")


magpie_data <- as.data.frame(table(d$magpie_class))

colnames(magpie_data) <- c("Magpie class", "N observations" )

library(knitr)
library(tidyr)


magpie_reshape <- select(d, magpie_class, agricultural_system)
magpie_reshape <- as.data.frame(table(magpie_reshape))

magpie_reshape <- magpie_reshape %>% group_by(magpie_class) %>% pivot_wider(names_from = agricultural_system,
                                          values_from = Freq)
```

```{r, echo = FALSE}

kable(magpie_data, row.names = TRUE)
kable(magpie_reshape, row.names = TRUE)

```


## For each crop, number of observations and papers 

For VividEconomics focus, the least data there is is on fruits and vegetables. 
This is because in the scoping phase, there wasn't a focus on these commodities.

Solution: provide a time estimate on how long it would take to get this data.

```{r, echo = FALSE}

# 1. Add magpie classes column

magpie <- read.csv2(here("Data/01.Processed_Data/04.Magpie_classes", "CSV_Magpie_classess.csv"))

crop_data <- as.data.frame(table(d$Crop))

  # Add column with magpie classes
colnames(crop_data)[1] <- "Crop"
colnames(crop_data)[2] <- "N"

colnames(magpie)[3] <- "My_crops"  # Change column name that bothered me

#Input magpie classes into my d data set
magpie <- na.omit(magpie)  # First I remove rows with NA in the magpie data set
                           # for two reasons: 
                           # 1. NA rows means I don't have data for that crop
                           # 2. NA's will interrupt the following for loop

crop_data$Magpie_class <- ""  # Create column where I will store the info

for (i in (1:10)){
  #browser()
  for (j in (1:31)){
    if (crop_data$Crop[i] == magpie$My_crops[j]){
      crop_data$Magpie_class[i] <- magpie$Magpie_classess[j]
    }}}

crop_data[c(1,3,10), 3] <- "Vegetables"
crop_data[8, 3] <- "Fruit"

crop_data <- relocate(crop_data, Magpie_class, .before = Crop) # Reorder columns

# 2. Count papers per crop

papers <- select(d, Crop, Paper_ID)

data_count_1 <- aggregate(data = papers,                  # Count the number of 
                          Paper_ID ~ Crop,                # per crop
                          function(x) length(unique(x)))

crop_Nstudies_Npapers <- merge(crop_data,data_count_1, by="Crop")  # Make table

    #Change colnames and oder in descending number of observations

colnames(crop_Nstudies_Npapers)[c(3,4)] <- c("N", "Number of papers" )

crop_Nstudies_Npapers <- crop_Nstudies_Npapers[order(-crop_Nstudies_Npapers$N),]


# 3. Make table

kable(crop_Nstudies_Npapers, row.names = TRUE)

```

## For each crop, number of observations per agricultural system

If transform all data into the same format, I could compare:

  - Maize: conservation, mixed, transgenic
  - Cotton: mixed and transgenic
  - Oil palm: all data on conventional
  - Soybean: not much, can get an estimate for mixed cropping
  - Sugarcane: conservation and conventional
  - Potato (potentially): conventional, organic, transgenic

```{r, echo = FALSE}

# Magpie reshape 

magpie_reshape <- select(d, Crop, agricultural_system)
magpie_reshape <- as.data.frame(table(magpie_reshape))

magpie_reshape <- magpie_reshape %>% group_by(Crop) %>% pivot_wider(names_from = agricultural_system,
                                          values_from = Freq)
kable(magpie_reshape, row.names = TRUE)

```

## For each crop, number of effect size observations

```{r, echo = FALSE}

library(tidyr)

d$Effect_size[d$Effect_size == "Hedges’ d"]<- "Hedges' d"
d$Effect_size[d$Effect_size == "Hedges_d"]<- "Hedges' d"

reshape <- select(d, Crop, Effect_size)
reshape <- as.data.frame(table(reshape))

reshape <- reshape %>% group_by(Crop) %>% pivot_wider(names_from = Effect_size,
                                          values_from = Freq)

kable(reshape, row.names = TRUE)


```


## For each prodcution system, number of observations and papers  

Potentially useful to do an analysis per agricultural system. 
```{r, echo = FALSE}

systems <- aggregate(data = d,                        # Count the number of papers
               Paper_ID ~ agricultural_system,        # per agricultural system
               function(x) length(unique(x)))                                     
               

systems_obs <- as.data.frame(table(d$agricultural_system))
colnames(systems_obs) <- c("agricultural_system", "N of observations")

systems <- merge( systems_obs, systems, by = "agricultural_system")
systems <- systems[order(-systems$`N of observations`),]

colnames(systems) <- c("agricultural_system", "N of observations", "N of papers")


kable(systems, row.names = TRUE)

```

## Conclusion

If all the data is transformed into the same format, some estimates can be obtained 
for all crops VividEconomics are interested in except for vegetables, fruits and 
nuts. Potato could potentially be analysed too, although I would recommend getting 
some a couple more papers on it. 

  1. Problem: since the data came from syntheses, many observations do not come with 
variance data, which is needed to perform a meta analysis. 

  - Solution: extract the data from the individual studies included in the synthesis.
However, that will postpone the meta analysis.

  - Solution: look for information on how to perform meta analysis without variance data.
  
  2. Problem: there is not enough data on fruits, vegetables and nuts to perform a 
meta analysis, since these were not focal crops in the scoping phase. 

  - Solution: provide time estimate on how long it would take to obtain this data.
  




