---
title: "23.Analysis_LRR"
author: "Christina Raw"
date: "28/3/2022"
output: html_document
---
In this script I am going to analyse biodiversity response to agricultural practices using the log response ratio data. This data comes either directly from log response ratio results of meta-analyses, or from back transforming percentage change data into log response ratio. 

```{r, include = FALSE}
rm(list=ls())
gc()

library(readxl)
library(here)
library(ggplot2)
library(lme4)
library(fitdistrplus)
library(blme)

d <- read_xlsx(here("Datasets", "06.Excel_Dataset_to_model_LRR.xlsx"))

```

## Distribution

As you can see, the data has a zero-inflated Gaussian distribution. This is die to two reasons:
  - There were 200 cases where biodiversity did not differ between the control and the treatment agricultural practices
  - When transforming from wide to long, I added the 349 LRR control values (which were all 0) to the LRR column
  
Therefore, there are 200+ 349 = 549 zeroes in the data set. In total, 78.65% of the data is 0

```{r, echo = FALSE}

hist(d$LRR, xlim = c(-4.477337, 2.586478), xlab = "LRR", main = "Histogram of LRR" )

descdist(d$LRR, discrete = FALSE)

qqnorm(d$LRR)

```

```{r, include = FALSE}

100*sum(d$LRR == 0)/nrow(d)

```

It seems like a zero-inflated normal distribution.

# Options I epxlored

## Option 1 = heavy-tailed distribution

## Try the package heavy, for heavy tailed distribution

It does not work because there aren't enough observations for certain groups. However, I don't think the problem is heavy tailed, but zero inflation.

```{r}
library(heavy)

d <-d[!is.na(d$biodiveristy_metric_category),]

m <- heavyLme(LRR ~ agricultural_system, random = ~ Crop, groups = ~ biodiveristy_metric_category,  data = d)
summary(m)

```

## Option 2: Bayesian approach

All the information I have found on modelling zero-inflated data with bayesian approach is limited to count data (GLMM)

    ## Blmer package

```{r}

m <- blmer(LRR ~ agricultural_system + magpie_class + biodiveristy_metric_category + (1|Crop) + (1|ID), data = d)

summary(m)
plot(m)

m <-blmer(formula, data = NULL, REML = TRUE,
      control = lmerControl(), start = NULL, verbose = 0L,
      subset, weights, na.action, offset, contrasts = NULL,
      devFunOnly = FALSE, cov.prior = wishart,
      fixef.prior = NULL, resid.prior = NULL, ...)


names(d)
```

    ## Pabkage glmTMB for zero-inflated generalized linear mixed modeling
    
# Option 3: hurdle models 

## Zero-inflated Gaussian mixed models(ZIGMMs) Xinyan Zhang, Boyi Guo, Nengjun Yi

These authors developed this model to overcome the limitations encountered in other models: zero-inflation, proportion data , missing data, time-dependency and correlations. 

We develop an efficient Expectation-Maximization (EM) algorithm to fit the ZIGMMs
by taking advantage of the standard procedure for fitting LMMs. We show computational efficiency of ZIGMMs compared with the other two zero-inflated methods, ZIBR and zeroinflated negative binomial mixed models implemented in the R package glmmTMB. Extensive
simulations demonstrate that our ZIGMMs outperform the various previously used methods in detecting associated effects in longitudinal microbiome data. We also apply our method to a
shotgun longitudinal microbiome proportion data.

For data with excessive zeros, it can be assumed that transformed values yij may come from either a
degenerate distribution having the point mass at zero (zero state) or a Gaussian (i.e., normal)
distribution [17]. Thus, the transformed values yij can be modeled with the zero-inflated
Gaussian distribution:

## More info on  how to model non-negative zero-inflated continuous data?

https://stats.stackexchange.com/questions/187824/how-to-model-non-negative-zero-inflated-continuous-data


# Option 4: linear model?

## Simple linear model to check it lm is a good model fit

By comparing estimated and predicted R2 it looks like a good fot. However, the model diagnostic plots indicate that this is not the best fit. Especially, the Q-Q plot indicates that I need a model for a non-normal distribution. This is bec 

```{r}

m <- lm(LRR ~ Treatment, data = d )
summary(m)
plot(m)

m1 <- lmer(LRR ~ Treatment + magpie_class + biodiveristy_metric_category + (1|Crop) + (1|ID), data = d )

summary(m1)
plot(m1)

m2 <- lmer(LRR ~ 1 + (1|Treatment), data = d)
summary(m2)

unique(d$Treatment)

treatment <- subset(d, Treatment != "Control")
qqnorm(d$LRR)

```

# Conclusion

The zero inflation an artifact of the data set structure, caused by adding all the control data which is = 0. The zero inflation is not actually biologically caused, it is only a consequence of the data set structure. Therefore, the hurdle model does not make sense biologically (remember, always chose model that makes biological sense). If it were, for example, presence (=1) absence (= 0) data, then a hurdle model would make sense biologically. Im my case, the zeroes are not true zeroes, true data.

Furthermore, the zero inflation should not influence the output. Imagine a random value is added to all the data, then the zero-inflation problem would be solved, but the model output would still remain the same. The zero-inflation is not actually causing any deviations to the results. 

# Next steps

Analysis with the Bayesian approach Andy suggested.




