---
title: "24.Analysis_LRR"
author: "Christina Raw"
date: "1/4/2022"
output: html_document
---  

Very preliminary analysis. Wheights are missing, have not looked at other formula structures or even reflected on the output yet. A haven't checked diagnostic plots either. Basically I have just written this and nothing else.

```{r, warning = FALSE, message = FALSE}

library(blme) # For Bayesian analysis
library(readxl) # For reading the excel data set
library(here) # For specifying file paths
library(knitr) # For making tables
library(kableExtra)# for pretty rmd tables
library(dplyr) # For some data wrangling
library(tidyr) # For some data wrangling

```

```{r, echo = FALSE}

rm(list=ls())
d <- read_excel(here("Datasets", "07.Excel_Dataset_to_model_LRR_LONG.xlsx"))

treatment_obs <- as.data.frame(table(d$Treatment))
colnames(treatment_obs) <- c("Treatment", "N_Observations")
treatment_obs <- treatment_obs[order(treatment_obs$N_Observations, decreasing = TRUE),]

kbl(treatment_obs, row.names = FALSE)

```

#

  - Collapse sustainable and organic, as organic is a sistainable agricultural practice. 
  - Include in the analysis only those treatments that have > 10 observations
  - Add weights

```{r, warning = FALSE, message = FALSE}
# Collapse organic and sustainable

d$Treatment[d$Treatment == "Organic"] <- "Sustainable"
treatment_obs <- subset(treatment_obs, treatment_obs$N_Observations >= 10)

# Subset agricultural systems with > 10 observations

d <- d[d$Treatment %in% treatment_obs$Treatment, ] 

```


```{r, warning = FALSE, message = FALSE, echo = FALSE}

x <- select(d, Synthesis_type, N_Studies, Weight) # Check adding weights it worked
x<- x[!duplicated(x),]

kbl(treatment_obs, row.names = FALSE, caption = "Agricultural systems with more than 10 observations")

```

# Analysis 

## Residuals histogram and qqplot of data **without the control zeroes**

It looks fairly normal to me!

```{r}

# Null model with only random effect to see qq plot of weighted residuals 

nonzero <- read_excel(here("Datasets", "06.Excel_Dataset_to_model_LRR_WIDE.xlsx"))

m0 <- blmer(LRR ~ 1 + (1|Crop), weights = Weight, data = nonzero)

hist(residuals(m0))
qqnorm(resid(m0))

```

## Residuals histogram and qqplot of data **with the control zeroes**

Zero-inflated, but that's all. The zeros won't influence in the output. 

```{r}

# Null model with only random effect to see qq plot of weighted residuals 

d <- read_excel(here("Datasets", "07.Excel_Dataset_to_model_LRR_LONG.xlsx"))

m1 <- blmer(LRR ~ 1 + (1|Crop), weights = Weight, data = d)

hist(residuals(m1))
qqnorm(resid(m1))
```
```{r, warning = FALSE, message = FALSE}

# Collapse organic and sustainable

d$Treatment[d$Treatment == "Organic"] <- "Sustainable"
treatment_obs <- subset(treatment_obs, treatment_obs$N_Observations >= 10)

# Subset agricultural systems with > 10 observations

d <- d[d$Treatment %in% treatment_obs$Treatment, ] 

```

## Bayesian model

```{r}

m3 <- blmer(LRR ~ Treatment + magpie_class + biodiveristy_metric_category + (1|Crop) + (1|ID), weights = Weight, data = d)

summary(m3)
```

I read on internet a way to fix the warnings https://stats.stackexchange.com/questions/242109/model-failed-to-converge-warning-in-lmer

```{r}

m4 <- blmer(LRR ~ Treatment + magpie_class + biodiveristy_metric_category + (1|Crop) + (1|ID), weights = Weight, data = d, control = lmerControl(optimizer ="Nelder_Mead"))

summary(m4)
plot(m4)

```

### 7/4/22

From the book on bayesian analysis - linear regression (p. 235): "However, even in the non informative case, posteriori simulations are useful for predictive inference and model checking"

From LME4 package: "Finally, posterior predictive simulation (Gelman and Hill 2006) is a generally useful diagnostic
tool, adapted from Bayesian methods, for exploring model fit. Users pick some summary
statistic of interest. They then compute the summary statistic for an ensemble of simulations
(Section 5.2.8), and see where the observed data falls within the simulated distribution; if the observed data is extreme, we might conclude that the model is a poor representation of reality. For example, using the sleep study fit and choosing the interquartile range of the reaction times as the summary statistic:
> iqrvec <- sapply(simulate(fm1, 1000), IQR)
> obsval <- IQR(sleepstudy$Reaction)
> post.pred.p <- mean(obsval >= c(obsval, iqrvec)


# Next steps today

## 1. Check model fit without prior

BAD

```{r}
# Diagnostic plots

plot(m4, type = c("p", "smooth"))

plot(m4, sqrt(abs(resid(.))) ~ fitted(.),
type = c("p", "smooth"))

library(lattice)
qqmath(m4, id = 0.05)

# MCMC diagnostic
# To understand whether there were problems with saplings. In other words, how reliable are our samples and thus our inference?

install.packages("bayesplot")
library(bayesplot)

# Trace plot of Linear Regression, define variables to plot via pars flag
mcmc_trace(m4,  pars = c("b_Intercept", "b_std_age", "sigma"), 
           n_warmup = 500, facet_args = list(ncol = 2, labeller = label_parsed))
           

# ACF plot for Linear Regression, lags can be easily defined in the function

m4_posteriors <- m4$posterior
mcmc_acf(lin_reg, pars = c("b_Intercept", "b_std_age", "sigma"), 
         lags = 15)

```




## 2. Check model fit with prior

But wait, the options for priors for fixed effect make no sense because they are categorical. Maybe prioris for the residuals? BUT this is too much for my knowledge so I will input NO priors 

```{r}
m5 <- blmer(LRR ~ Treatment + magpie_class + biodiveristy_metric_category + (1|Crop) + (1|ID), weights = Weight, data = d, control = lmerControl(optimizer ="Nelder_Mead"),  fixef.prior = normal)

summary(m5)
plot(m5)
plot(m5, type = c("p", "smooth"))

plot(m5, sqrt(abs(resid(.))) ~ fitted(.),
type = c("p", "smooth"))

library(lattice)
qqmath(m5, id = 0.05)


```



Actually now I am thinking of doing two separate models. One for crops and one for magpie classes 


```{r}

# Crop
 
unique(d$magpie_class) # fixed, levels = 5
unique(d$biodiveristy_metric_category) # random (> 5 levels), also, because is clustering term, nor fized
unique(d$Crop)

m4 <- blmer(LRR ~ Treatment + magpie_class + biodiveristy_metric_category + (1|Crop) + (1|ID), weights = Weight, data = d, control = lmerControl(optimizer ="Nelder_Mead"))

summary(m4)

mCrop <- blmer(LRR ~ Treatment + (1|biodiveristy_metric_category) + (1|Crop) + (1|ID), weights = Weight, data = d, control = lmerControl(optimizer ="Nelder_Mead"))
summary(mCrop)

```




