---
title: "25.Analysis_LRR"
author: "Christina Raw"
date: "8/4/2022"
output: html_document
---

In this script I am going to perform a Bayesian analysis of the data.

  - Data set: quantitative observation of biodiversity under different agricultural systems. 
  - Goal: see how different agricultural systems impact biodiversity.

## Brief introduction of the method

I will perform a linear mixed model in a Bayesian environment implementing the methods of Chung, et al. (2013). 

**Chung *et al.* (2013)** *what I have understood*

When there is a case of small number of groups in Linear Mixed Models, estimates of group-level variance (i.e., between-group variance) can be noisy and often be 0, when, in reality, *there is* between group variance. In this case, accepting the null hypothesis that group-level variance is 0 is a wrong assumption that can lead to wrong conclusions. Chung *et al.* solve this problem by applying a Maximum Penalized Likelihood (MPL) estimator for variance, which adds a penalization of +1 to variance when is estimated as 0. In this case:

  - The exponential of the penalty can be regarded as a Bayesian prior density for variance
  - The MPL estimates can be regarded as posterior modal estimates
  - The MPL is as efficient as the MLE
  - It is possible to apply the MPL without requiring strong prior knowledge
  - The MPL does not involve simulation so there are no MCMC iterations (I think)

```{r, warning = FALSE, message = FALSE}

library(blme) # For Bayesian analysis
library(readxl) # For reading the excel data set
library(here) # For specifying file paths
library(knitr) # For making tables
library(kableExtra)# for pretty rmd tables
library(dplyr) # For some data wrangling
library(tidyr) # For some data wrangling

```

## 1. Take a look at the data

```{r, echo = FALSE}

d <- read_excel(here("Datasets", "07.Excel_Dataset_to_model_LRR_LONG.xlsx"))

treatment_obs <- as.data.frame(table(d$Treatment))
colnames(treatment_obs) <- c("Treatment", "N_Observations")
treatment_obs <- treatment_obs[order(treatment_obs$N_Observations, decreasing = TRUE),]

kbl(treatment_obs, row.names = FALSE)

```

#

  - Collapse sustainable and organic, as organic is a sustainable agricultural practice. 
  - Include in the analysis only those treatments that have > 10 observations
  - Add weights

```{r, warning = FALSE, message = FALSE}

# Collapse organic and sustainable

d$Treatment[d$Treatment == "Organic"] <- "Sustainable"

treatment_obs <- as.data.frame(table(d$Treatment))
colnames(treatment_obs) <- c("Treatment", "N_Observations")
treatment_obs <- treatment_obs[order(treatment_obs$N_Observations, decreasing = TRUE),]

treatment_obs <- subset(treatment_obs, treatment_obs$N_Observations >= 10)

# Subset agricultural systems with >= 10 observations

d <- d[d$Treatment %in% treatment_obs$Treatment, ] 

```

### Residuals histogram and qqplot of data **without the control zeroes**

It looks fairly normal to me!

```{r}

# Null model with only random effect to see qq plot of weighted residuals 

nonzero <- read_excel(here("Datasets", "06.Excel_Dataset_to_model_LRR_WIDE.xlsx"))

m0 <- blmer(LRR ~ 1 + (1|Crop), weights = Weight, data = nonzero)

hist(residuals(m0))
qqnorm(resid(m0))

```

### Residuals histogram and qqplot of data **with the control zeroes**

Zero-inflated, but that's all. The zeros won't influence in the output. 

```{r}

# Null model with only random effect to see qq plot of weighted residuals 

m1 <- blmer(LRR ~ 1 + (1|Crop), weights = Weight, data = d)

hist(residuals(m1))
qqnorm(resid(m1))
```

## 2. Analysis

### Relevel d so conventional is reference

```{r}
d$Treatment <- as.factor(d$Treatment)
d$Treatment <- relevel(d$Treatment, ref = "Conventional")
```

### Model 1: how does agriculture vary across agricultural systems?

```{r}

m1 <- blmer(LRR ~ Treatment + biodiveristy_metric_category + (1|ID) + (1|Crop), data = d, control = lmerControl(optimizer ="Nelder_Mead"))

summary(m1)
```

Crop does not explain much variance as random effect. There is significant biodiversity change between conventional and: conservation, primary vegetation and sustainable (organic + sustainable).

However, this does not tell me how each metric varies under the different agricultural systems. To see that, I would need to include an interaction (*I think*).

### Model 2: how does agriculture vary across agricultural systems with a closer look to biodiversity metric.

  - Fixed effects (I am interested in their mean)
    - Agricultural system: I want to know biodiversity mean in each agricultural system ( = Treatment)
    - Biodiversity metric: I want to know the mean of each type of biodiversity metric **in each type of agricultural system**, so, maybe introduce an interaction? metric:agricultural system?

  - Random effects (variance)
    - Crop: grouping factor
    - ID: actually this one is only to account for non independence of the two data points that come from the same control-treatment comparison

```{r}

m2 <- blmer(LRR ~  Treatment + biodiveristy_metric_category:Treatment + (1|Crop) + (1|ID), weights = Weight, data = d, control = lmerControl(optimizer ="Nelder_Mead"))

summary(m2)

```

Error occurs because not all combinations of the interaction have data, so the model drops them.

Residual variance increases when including the interaction, which I think means that this model is worse explaining the variance.

Under mixed agriculture, both diversity and development differ significantly from intercept.

**HOWEVER**

I am not really interested in how the metrics vary per system. My *goal* with this analysis is too see how biodiversity varies under agricultural systems **whithin crop and magpie class**. Therefore, I believe subsetting first per crop and then per magpie class will be more inormative.

### Analysis per crop

  - 1) Subset per crop
  
```{r, include = FALSE}
crop <- unique(d$Crop)

crop_list <- list() 

for (i in crop){
data <- dplyr :: filter( d , Crop == i)
crop_list[[i]] <- data
}

list2env(lapply(crop_list, as.data.frame), .GlobalEnv)

```

  - 2) Do model for crops with more than 50 observations (rule of thumb for LMM)
  
```{r}
# Cassava: only two treatments (mixed and monoculture) -> t.test

t.test(LRR~Treatment, data = Cassava )

# Cotton

cotton <- blmer(LRR ~  Treatment + biodiveristy_metric_category:Treatment  + (1|ID), weights = Weight, data = Cotton, control = lmerControl(optimizer ="Nelder_Mead"))

summary(cotton)

# Maize

maize <- blmer(LRR ~  Treatment + biodiveristy_metric_category:Treatment  + (1|ID), weights = Weight, data = Cotton, control = lmerControl(optimizer ="Nelder_Mead"))

summary(maize)
```

Already I can see that doing it per crop is not the best approach given the data set. Next, I will try per magpie class.

### Analysis per magpie class

  - 1) Subset per magpie class
  
```{r, include = FALSE}
magpie <- unique(d$magpie_class)

magpie_list <- list() 

for (i in magpie){
data <- dplyr :: filter( d , magpie_class == i)
magpie_list[[i]] <- data
}

list2env(lapply(magpie_list, as.data.frame), .GlobalEnv)

```

  - 2) Analysis per magpie class
  
```{r}
cereals <- blmer(LRR ~  Treatment + biodiveristy_metric_category:Treatment  + (1|ID), weights = Weight, data = Cereals, control = lmerControl(optimizer ="Nelder_Mead"))

summary(cereals)
  
```


```{r}
oil <- blmer(LRR ~  Treatment + biodiveristy_metric_category:Treatment  + (1|ID), weights = Weight, data = `Oil crops`, control = lmerControl(optimizer ="Nelder_Mead"))

summary(oil)

```
I could continue but I guess I am going to keep getting these really small estimates. 

I think what is happening is that when i subset per crop and magpie class, the zeroes included in the control are playing a much stronger role and actually *do* influence the model output.

Comments?





