---
title: "26.Analysis_LRR"
author: "Christina Raw"
date: "19/4/2022"
output: html_document
---

Table of contents:

  - 1. Take a look at the data
    - Residuals histogram and qqplot of data **without the control zeroes**
    - Residuals and qqplot of data **with the control zeroes**
  - 2. Analysis
    - 2.1 Re level data so conventional is reference
    - 2.2 Crop as random effect
    - 2.3 Crop as fixed effect
    - 2.4 Biodiversity metric in each system
    - 2.5 Model selection
    - 2.6 Model goodness-of-fit
  - 3. Analysis per biodiversity metric
    - 3.1 Biodiversity metric data barplot
    - 3.2 Biodiversity metric data **per crop**
    - 3.3 Biomass and diversity analysis **per crop**: Maize and Oil palm
    - 3.4 Biodiversity metric data **per magpie class**
    - 3.5 biomass, diversity and enzymatic activity analysis **per magpie class**: Cereals and Oil crops

```{r, warning = FALSE, message = FALSE}

library(blme) # For Bayesian analysis
library(readxl) # For reading the excel data set
library(here) # For specifying file paths
library(knitr) # For making tables
library(kableExtra)# for pretty rmd tables
library(dplyr) # For some data wrangling
library(tidyr) # For some data wrangling
library(ggplot2) # For making cool plots
library(MuMIn) # For model selection

```

After my meeting with Andy on 14/4/22, we decided to input crop / magpie class as fixed effect. Also, I will see what biodiversity metric looks like and subset by categories, then rerun model -\> see estimates of biodiversity metric categories in space and see how differently they vary under agricultural systems.

## 1. Take a look at the data

```{r, include = FALSE}

d <- read_excel(here("Datasets", "07.Excel_Dataset_to_model_LRR_LONG.xlsx"))

str(d)
d$Crop <- as.factor(d$Crop)
d$Treatment <- as.factor(d$Treatment)
d$magpie_class <- as.factor(d$magpie_class)
d$Weight <- as.numeric(d$Weight)
d$biodiveristy_metric_category <- as.factor(d$biodiveristy_metric_category)

```

```{r, echo = FALSE}

treatment_obs <- as.data.frame(table(d$Treatment))
colnames(treatment_obs) <- c("Treatment", "N_Observations")
treatment_obs <- treatment_obs[order(treatment_obs$N_Observations, decreasing = TRUE),]

kbl(treatment_obs, row.names = FALSE)

```

# 

-   Collapse sustainable and organic, as organic is a sustainable agricultural practice.
-   Include in the analysis only those treatments that have \> 10 observations
-   Add weights

```{r, warning = FALSE, message = FALSE}

# Collapse organic and sustainable

d$Treatment[d$Treatment == "Organic"] <- "Sustainable"

treatment_obs <- as.data.frame(table(d$Treatment))
colnames(treatment_obs) <- c("Treatment", "N_Observations")
treatment_obs <- treatment_obs[order(treatment_obs$N_Observations, decreasing = TRUE),]

treatment_obs <- subset(treatment_obs, treatment_obs$N_Observations >= 10)

# Subset agricultural systems with >= 10 observations

d <- d[d$Treatment %in% treatment_obs$Treatment, ] 

```

### Residuals histogram and qqplot of data **without the control zeroes**

It looks fairly normal to me!

```{r}

# Null model with only random effect to see qq plot of weighted residuals 

nonzero <- read_excel(here("Datasets", "06.Excel_Dataset_to_model_LRR_WIDE.xlsx"))

m0 <- blmer(LRR ~ 1 + (1|Crop), weights = Weight, data = nonzero)

hist(residuals(m0))
qqnorm(resid(m0))

```

### Residuals histogram and qqplot of data **with the control zeroes**

Zero-inflated, but that's all. The zeros won't influence in the output.

```{r}

# Null model with only random effect to see qq plot of weighted residuals 

m1 <- blmer(LRR ~ 1 + (1|Crop), weights = Weight, data = d)

hist(residuals(m1))
qqnorm(resid(m1))
```

## 2. Analysis

### 2.1 Relevel d so conventional is reference

```{r}
d$Treatment <- as.factor(d$Treatment)
d$Treatment <- relevel(d$Treatment, ref = "Conventional")
```

### 2.2 Crop as random effect

```{r, error=FALSE, message = FALSE}
m0 <- blmer(LRR ~ Treatment + magpie_class + biodiveristy_metric_category + (1|Crop) + (1|ID), weights = Weight, data = d, control = lmerControl(optimizer ="Nelder_Mead"))

summary(m0)
```

### 2.3 Crop as fixed effect

```{r, error=FALSE, message = FALSE}
m1 <- blmer(LRR ~ Treatment + magpie_class + biodiveristy_metric_category + Crop + (1|ID), weights = Weight, data = d, control = lmerControl(optimizer ="Nelder_Mead"))

summary(m1)
```

### 2.4 Biodiversity metric in each system

I want to know the mean of each type of biodiversity metric in each type of agricultural system so, maybe I am going to introduce the interaction -\> metric:agricultural system

```{r, message = FALSE}
m2 <- blmer(LRR ~  Treatment + biodiveristy_metric_category:Treatment + (1|Crop) + (1|ID), weights = Weight, data = d, control = lmerControl(optimizer ="Nelder_Mead"))

summary(m2)
```

### 2.5 Model selection

```{r}

anova(m0, m1, m2)

```
According to the anova output, the model that includes the interaction (m3) is a significantly better fit to the data. **However**, I am not sure I can use this anova because blmer uses a penalised maximum likelihood estimator of variance, whereas for doing the anova, a maximum likelihood estimator is needed, loosing the whole point of blmer.

```{r, warning = FALSE}
r.squaredGLMM(m0)
r.squaredGLMM(m1)
r.squaredGLMM(m2)

model.sel(m0,m1,m2)

```

Model 2 seems like the best fit: highest r2, lowest logLik, lowest AIC.

### 2.6 Model goodness-of-fit

```{r}

# Residuals vs Fitted

plot(m2)

# Normality of residuals

qqnorm(residuals(m2))

# Residuals vs Leverage

ggplot(data.frame( lev = hatvalues(m2), pearson = residuals( m2, type = "pearson")),
      aes(x = lev, y = pearson)) +
    geom_point() +
    theme_bw()

```


```{r}
## scale-location plot, with red smoothed line

scale_loc_plot <- function(m, line.col = "red", line.lty = 1,
                           line.lwd = 2) {
  plot(m0, sqrt(abs(resid(.))) ~ fitted(.),
       type = c("p", "smooth"),
       par.settings = list(plot.line =
                             list(alpha=1, col = line.col,
                                  lty = line.lty, lwd = line.lwd)))
}
scale_loc_plot(m0)
```



## 3. Analysis per biodiversity metric

### 3.1 Biodiversity metric data barplot

```{r, include = FALSE}
# Make separate data frames for each biodiversity metric category

metric <- unique(d$biodiveristy_metric_category)

metric_list <- list() 

for (i in metric){
data <- dplyr :: filter( d , biodiveristy_metric_category == i)
metric_list[[i]] <- data
}

list2env(lapply(metric_list, as.data.frame), .GlobalEnv)

```

```{r, echo = FALSE}

ggplot(d, aes(x = biodiveristy_metric_category, fill = agricultural_system)) +
              geom_bar() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### 3.2 Biodiversity metric data per **crop**

An analysis could be performed for: biomass (maize) and diversity (oil palm)

```{r, echo = FALSE}

# Biomass

kable(table(biomass$Crop, biomass$agricultural_system), caption = "Metric: biomass") %>%
  kable_styling(full_width = F, position = "float_left")

# Development

kable(table(development$Crop, development$agricultural_system), caption = "Metric: development") %>%
  kable_styling(full_width = F, position = "right")

# Diversity

kable(table(diversity$Crop, diversity$agricultural_system), caption = "Metric: diversity")

# Efficiency

kable(table(efficiency$Crop, efficiency$agricultural_system), caption = "Metric: efficiency") %>%
  kable_styling( full_width = F, position = "float_left")

# Enzymatic activity 

kable(table(Enzymatic_activity$Crop, Enzymatic_activity$agricultural_system), caption = "Metric: Enzymatic_activity") %>%
  kable_styling( full_width = F, position = "right")

# Reproduction

kable(table(reproduction$Crop, reproduction$agricultural_system), caption = "Metric: reproduction") %>%
  kable_styling( full_width = F, position = "float_left")

# Survival

kable(table(survival$Crop, survival$agricultural_system), caption = "Metric: survival") %>%
  kable_styling( full_width = F, position = "right")

```

### 3.3 Biomass and diversity analysis per **crop**: Maize and Oil Palm

An analysis could be performed for: biomass (maize) and diversity (oil palm)

```{r}
# Biomass: Maize

maize <- subset(d, Crop == "Maize" & biodiveristy_metric_category == "biomass")

unique(maize$Paper_ID)
```

There are three papers that provide biomass data fro maize crop

```{r}
model_maize <- blmer(LRR ~  Treatment + (1|ID), weights = Weight, data = maize, control = lmerControl(optimizer ="Nelder_Mead"))

summary(model_maize)
```

```{r}

# Diversity: Oil Palm

oil <- subset(d, Crop == "Oil_palm" & biodiveristy_metric_category == "diversity")

unique(oil$Paper_ID)
```
There are 6 papers that provide diversity data for oil palm crop

```{r}
model_oil <- blmer(LRR ~  Treatment + (1|ID), weights = Weight, data = oil, control = lmerControl(optimizer ="Nelder_Mead"))

summary(model_oil) 

```

### 3.4 Biodiversity metric data per **magpie class**

An analysis could be performed for biomass (cereals and oil crops), diversity (oil crops) and enzymatic activity (oil crops)

```{r, echo = FALSE}

# Biomass

kable(table(biomass$magpie_class, biomass$agricultural_system), caption = "Metric: biomass")  %>%
  kable_styling( full_width = F, position = "float_left") 

# Development

kable(table(development$magpie_class, development$agricultural_system), caption = "Metric: development") %>%
  kable_styling( full_width = F, position = "right")

# Diversity

kable(table(diversity$magpie_class, diversity$agricultural_system), caption = "Metric: diversity") 

# Efficiency
  
kable(table(efficiency$magpie_class, efficiency$agricultural_system), caption = "Metric: efficiency") %>%
  kable_styling( full_width = F, position = "float_left")

# Enzymatic activity

kable(table(Enzymatic_activity$magpie_class, Enzymatic_activity$agricultural_system), caption = "Metric: Enzymatic_activity") %>%
  kable_styling( full_width = F, position = "right")

# Reproduction

kable(table(reproduction$magpie_class, reproduction$agricultural_system), caption = "Metric: reproduction") %>%
  kable_styling( full_width = F, position = "float_left")

# Survival

kable(table(survival$magpie_class, survival$agricultural_system), caption = "Metric: survival") %>%
  kable_styling( full_width = F, position = "right")

```

### 3.5 Biomass, diversity and enzymatic activity analysis per **magpie class**: cereals and oil crops

An analysis could be performed for biomass (cereals and oil crops), diversity (oil crops) and enzymatic activity (oil crops)

```{r}
# Biomass: Cereals 

cereals <- subset(d, magpie_class == "Cereals" & biodiveristy_metric_category == "biomass")

unique(cereals$Paper_ID)

```
There are four papers that provide data for biomass for the cereals magpie class

```{r}

biomass_cereals <- blmer(LRR ~  Treatment + (1|ID), weights = Weight, data = cereals, control = lmerControl(optimizer ="Nelder_Mead"))

summary(biomass_cereals)

```

```{r}
# Biomass: Oil crops

oil_crops <- subset(d, magpie_class == "Oil crops" & biodiveristy_metric_category == "biomass")

unique(oil_crops$Paper_ID)
```

There are two papers that provide data for biomass for the oil crops magpie class

```{r}

biomass_oil_crops <- blmer(LRR ~  Treatment + (1|ID), weights = Weight, data = oil_crops, control = lmerControl(optimizer ="Nelder_Mead"))

summary(biomass_oil_crops)

```

```{r}
# Diversity: Oil crops

oil_crops <- subset(d, magpie_class == "Oil crops" & biodiveristy_metric_category == "diversity")

unique(oil_crops$Paper_ID)
```

There are nine papers that provide data for diversity for the oil crops magpie class

```{r}

diversity_oil_crops <- blmer(LRR ~  Treatment + (1|ID), weights = Weight, data = oil_crops, control = lmerControl(optimizer ="Nelder_Mead"))

summary(diversity_oil_crops)

```
```{r}
# Enzymatic activity: Oil crops

oil_crops <- subset(d, magpie_class == "Oil crops" & biodiveristy_metric_category == "Enzymatic_activity")

unique(oil_crops$Paper_ID)

```

There are two papers that provide enzymatic activity data for the oil crops magpie class

```{r}

enzymatic_oil_crops <- blmer(LRR ~  Treatment + (1|ID), weights = Weight, data = oil_crops, control = lmerControl(optimizer ="Nelder_Mead"))

summary(enzymatic_oil_crops)

```

