---
title: "26.Analysis_LRR"
author: "Christina Raw"
date: "19/4/2022"
output: html_document
---

In this script I am going to perform a Bayesian analysis of the data to see how differet agricultural systems impact biodiversity. The data set consists of quantitative observation of biodiversity under different agricultural systems. 

Table of contents:

1. Take a look at the data
  - 1.1 Residuals histogram and qqplot of data **without the control zeroes**
  - 1.2 Residuals and qqplot of data **with the control zeroes**
2. Analysis
  - 2.1 Brief introduction of the method
  - 2.2 Re level data so conventional is reference
  - 2.3 Fixed effects
  - 2.4 Interaction Treatment:Biodiversity_metric_category
  - 2.5 Random effects
  - 2.6 Check R^2^
  - 2.7 Model
  - 2.8 Model goodness-of-fit
  - 2.9 Model with Crop as fixed effect
3. Analysis per biodiversity metric
  - 3.1 Biodiversity metric data barplot
  - 3.2 Biodiversity metric data **per crop**
  - 3.3 Biomass and diversity analysis **per crop**: Maize and Oil palm
  - 3.4 Biodiversity metric data **per magpie class**
  - 3.5 biomass, diversity and enzymatic activity analysis **per magpie class**: Cereals and Oil crops

```{r, warning = FALSE, message = FALSE}

library(blme) # For Bayesian analysis
library(readxl) # For reading the excel data set
library(here) # For specifying file paths
library(knitr) # For making tables
library(kableExtra)# for pretty rmd tables
library(dplyr) # For some data wrangling
library(tidyr) # For some data wrangling
library(ggplot2) # For making cool plots
library(MuMIn) # For model selection

```

After my meeting with Andy on 14/4/22, we decided to input crop / magpie class as fixed effect. Also, I will see what biodiversity metric looks like and subset by categories, then rerun model and see estimates of biodiversity metric categories in space and see how differently they vary under agricultural systems.

## 1. Take a look at the data

```{r, include = FALSE}

d <- read_excel(here("Datasets", "07.Excel_Dataset_to_model_LRR_LONG.xlsx"))

str(d)
d$Crop <- as.factor(d$Crop)
d$Treatment <- as.factor(d$Treatment)
d$magpie_class <- as.factor(d$magpie_class)
d$Weight <- as.numeric(d$Weight)
d$biodiveristy_metric_category <- as.factor(d$biodiveristy_metric_category)

```

```{r, echo = FALSE}

treatment_obs <- as.data.frame(table(d$Treatment))
colnames(treatment_obs) <- c("Treatment", "N_Observations")
treatment_obs <- treatment_obs[order(treatment_obs$N_Observations, decreasing = TRUE),]

kbl(treatment_obs, row.names = FALSE)

```



-   Collapse sustainable and organic, as organic is a sustainable agricultural practice.
-   Include in the analysis only those treatments that have \> 10 observations
-   Add weights

```{r, warning = FALSE, message = FALSE}

# Collapse organic and sustainable

d$Treatment[d$Treatment == "Organic"] <- "Sustainable"

treatment_obs <- as.data.frame(table(d$Treatment))
colnames(treatment_obs) <- c("Treatment", "N_Observations")
treatment_obs <- treatment_obs[order(treatment_obs$N_Observations, decreasing = TRUE),]

treatment_obs <- subset(treatment_obs, treatment_obs$N_Observations >= 10)

# Subset agricultural systems with >= 10 observations

d <- d[d$Treatment %in% treatment_obs$Treatment, ] 

```

### 1.1 Residuals histogram and qqplot of data **without the control zeroes**

It looks fairly normal to me!

```{r}

# Null model with only random effect to see qq plot of weighted residuals 

nonzero <- read_excel(here("Datasets", "06.Excel_Dataset_to_model_LRR_WIDE.xlsx"))

m0 <- blmer(LRR ~ 1 + (1|Crop), weights = Weight, data = nonzero)

hist(residuals(m0))
qqnorm(resid(m0))

```

### 1.2 Residuals histogram and qqplot of data **with the control zeroes**

Zero-inflated, but that's all. The zeros won't influence in the output.

```{r}

# Null model with only random effect to see qq plot of weighted residuals 

m1 <- blmer(LRR ~ 1 + (1|Crop), weights = Weight, data = d)

hist(residuals(m1))
qqnorm(resid(m1))
```

## 2. Analysis

### 2.1 Brief introduction of the method

I will perform a linear mixed model in a Bayesian environment implementing the methods of Chung, et al. (2013). 

**Chung *et al.* (2013)** *what I have understood*

When there is a case of small number of groups in Linear Mixed Models, estimates of group-level variance (i.e., between-group variance) can be noisy and often be 0, when, in reality, *there is* between group variance. In this case, accepting the null hypothesis that group-level variance is 0 is a wrong assumption that can lead to wrong conclusions. Chung *et al.* solve this problem by applying a Maximum Penalized Likelihood (MPL) estimator for variance, which adds a penalization of +1 to variance when is estimated as 0. In this case:

  - The exponential of the penalty can be regarded as a Bayesian prior density for variance
  - The MPL estimates can be regarded as posterior modal estimates
  - The MPL is as efficient as the MLE
  - It is possible to apply the MPL without requiring strong prior knowledge
  - The MPL does not involve simulation so there are no MCMC iterations (I think)

### 2.2 Relevel data set so conventional is the reference category

```{r}
d$Treatment <- as.factor(d$Treatment)
d$Treatment <- relevel(d$Treatment, ref = "Conventional")
```

### 2.3 Fixed effects

I want to include crop and treatment as fixed effect. I will construct a maximal model with them and do a round of model selection with the MuMim package.

```{r, error = FALSE, warning = FALSE, message = FALSE}

# Model

options(na.action = "na.fail")

m0 <- blmer(LRR ~ Treatment + Crop +(1|ID), weights = Weight, data = d, control = lmerControl(optimizer ="Nelder_Mead"))

# Model selection

dredge(m0, beta = "none", evaluate = TRUE)

```

According to the model selection table, including Crop as fixed effect is a worse model fit (lower AIC). Therefore, I will only inlcude Treatment as fixed effect.


### 2.4 Interaction treatment and biodiversity metric category

I want to know the mean of each type of biodiversity metric in each type of agricultural system so, I am going to introduce the interaction biodiversity_metric_category:Treatment, and see whether that is a better model fit.

```{r, message = FALSE, warning = FALSE, message = FALSE}

# Model selection

m0 <- blmer(LRR ~ Treatment +(1|ID), weights = Weight, data = d, control = lmerControl(optimizer ="Nelder_Mead"))

m1 <- blmer(LRR ~ Treatment + Crop +(1|ID), weights = Weight, data = d, control = lmerControl(optimizer ="Nelder_Mead"))

m2 <- blmer(LRR ~  Treatment + biodiveristy_metric_category:Treatment + (1|ID), weights = Weight, data = d, control = lmerControl(optimizer ="Nelder_Mead"), na.action = "na.omit")

m3 <- blmer(LRR ~  Treatment + Crop + biodiveristy_metric_category:Treatment + (1|ID), weights = Weight, data = d, control = lmerControl(optimizer ="Nelder_Mead"), na.action = "na.omit")

# Model selection. I am doing it this way instead of with the dredge function because with the dredge function I have to apply na.fail, and when I do so, I cannot run the model with the interaction. 

model.sel(m0, m1, m2, m3)

r.squaredGLMM(m0)
r.squaredGLMM(m1)
r.squaredGLMM(m2)
r.squaredGLMM(m3)

```

Including the interaction is a worse model fit, so I will not include it in the model.

### 2.5 Random effects

Since Crop is a clustering factor of the data, I want to see whether including crop as random effect is a better fit for the model.

```{r, warning = FALSE, message = FALSE}

# Model

m2 <- blmer(LRR ~  Treatment + (1|ID), weights = Weight, data = d, control = lmerControl(optimizer ="Nelder_Mead"), na.action = "na.omit")

m3 <- blmer(LRR ~  Treatment + (1|ID) + (1|Crop), weights = Weight, data = d, control = lmerControl(optimizer ="Nelder_Mead"), na.action = "na.omit")

# Model selection

anova(m2, m3)

```

Including crop as random effect yields basically the same model fit.

### 2.6 Check R^2^

I also want to check the R2.  

```{r, warning = FALSE, message = FALSE}
r.squaredGLMM(m0)
r.squaredGLMM(m1)
r.squaredGLMM(m2)
r.squaredGLMM(m3)
```
Again, the model with the interaction is a better fit and Crop as random effect does no make much difference. 

### 2.7 Model 

```{r}

m2 <- blmer(LRR ~  Treatment + (1|ID) + (1|Crop), weights = Weight, data = d, control = lmerControl(optimizer ="Nelder_Mead"), na.action = "na.omit")

summary(m2)

```

Compared to conventional agriculture, biodiversity was significantly higher under conservation agriculture ($\overline{U}_{LRR}$ = 0.49, CI: 0.40 to 0.57, t = 12.236). However, it was lower in disturbed forest ($\overline{U}_{LRR}$ = -0.86,  CI: 1.43 to -0.28, t = -2.775), primary vegetation ($\overline{U}_{LRR}$ = -0.7, CI: -1.01 to -0.37, t = -4.002) and sustainable agriculture ($\overline{U}_{LRR}$ = -2.290548, CI:-4.233562 to -0.3475343, t = -2.260)

These results are against what I would expect. 

### 2.8  Model goodness-of-fit

Now, I will assess the model's goodness. The model is model 2, wich includes the interaction biodiversity_metric_category:Treatment and Crop as random effect.

```{r}
# Residuals vs Fitted

plot(m2)
```

Non-linear relationship and Heteroscedasticity.

```{r}
# Normality of residuals

qqnorm(residuals(m2))
```

Actually not too bad, reflects zero inflation.

```{r}
# Residuals vs Leverage

ggplot(data.frame( lev = hatvalues(m2), pearson = residuals( m2, type = "pearson")),
      aes(x = lev, y = pearson)) +
    geom_point() +
    theme_bw()



```


```{r}
## scale-location plot, with red smoothed line

scale_loc_plot <- function(m, line.col = "red", line.lty = 1,
                           line.lwd = 2) {
  plot(m0, sqrt(abs(resid(.))) ~ fitted(.),
       type = c("p", "smooth"),
       par.settings = list(plot.line =
                             list(alpha=1, col = line.col,
                                  lty = line.lty, lwd = line.lwd)))
}

scale_loc_plot(m0)
```

### 2.9 Model with Crop as fixed effect

I will also show the results of the model with Crop as fixed effect, since it can also be of interest.

```{r}

m3 <- blmer(LRR ~ Treatment + Crop + biodiveristy_metric_category:Treatment + (1|ID), weights = Weight, data = d, control = lmerControl(optimizer ="Nelder_Mead"), na.action = "na.omit")

summary(m3)

```

### Taxon level

```{r}

m3 <- blmer(LRR ~ Treatment + Kingdom:Treatment + (1|ID), weights = Weight, data = d, control = lmerControl(optimizer ="Nelder_Mead"), na.action = "na.omit")

summary(m3)

sum(is.na(d$Phylum))
sum(is.na(d$Kingdom))

```


## 3. Analysis per biodiversity metric

### 3.1 Biodiversity metric data barplot

```{r, include = FALSE}
# Make separate data frames for each biodiversity metric category

metric <- unique(d$biodiveristy_metric_category)

metric_list <- list() 

for (i in metric){
data <- dplyr :: filter( d , biodiveristy_metric_category == i)
metric_list[[i]] <- data
}

list2env(lapply(metric_list, as.data.frame), .GlobalEnv)

```

```{r, echo = FALSE}

ggplot(d, aes(x = biodiveristy_metric_category, fill = agricultural_system)) +
              geom_bar() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### 3.2 Biodiversity metric data per **crop**

An analysis could be performed for: biomass (maize) and diversity (oil palm)

```{r, echo = FALSE}

# Biomass

kable(table(biomass$Crop, biomass$agricultural_system), caption = "Metric: biomass") %>%
  kable_styling(full_width = F, position = "float_left")

# Development

kable(table(development$Crop, development$agricultural_system), caption = "Metric: development") %>%
  kable_styling(full_width = F, position = "right")

# Diversity

kable(table(diversity$Crop, diversity$agricultural_system), caption = "Metric: diversity")

# Efficiency

kable(table(efficiency$Crop, efficiency$agricultural_system), caption = "Metric: efficiency") %>%
  kable_styling( full_width = F, position = "float_left")

# Enzymatic activity 

kable(table(Enzymatic_activity$Crop, Enzymatic_activity$agricultural_system), caption = "Metric: Enzymatic_activity") %>%
  kable_styling( full_width = F, position = "right")

# Reproduction

kable(table(reproduction$Crop, reproduction$agricultural_system), caption = "Metric: reproduction") %>%
  kable_styling( full_width = F, position = "float_left")

# Survival

kable(table(survival$Crop, survival$agricultural_system), caption = "Metric: survival") %>%
  kable_styling( full_width = F, position = "right")

```

### 3.3 Biomass and diversity analysis per **crop**: Maize and Oil Palm

An analysis could be performed for: biomass (maize) and diversity (oil palm)

```{r}
# Biomass: Maize

maize <- subset(d, Crop == "Maize" & biodiveristy_metric_category == "biomass")

unique(maize$Paper_ID)
```

There are three papers that provide biomass data fro maize crop

```{r}
model_maize <- blmer(LRR ~  Treatment + (1|ID), weights = Weight, data = maize, control = lmerControl(optimizer ="Nelder_Mead"))

summary(model_maize)
```

```{r}

# Diversity: Oil Palm

oil <- subset(d, Crop == "Oil_palm" & biodiveristy_metric_category == "diversity")

unique(oil$Paper_ID)
```
There are 6 papers that provide diversity data for oil palm crop

```{r}
model_oil <- blmer(LRR ~  Treatment + (1|ID), weights = Weight, data = oil, control = lmerControl(optimizer ="Nelder_Mead"))

summary(model_oil) 

```

### 3.4 Biodiversity metric data per **magpie class**

An analysis could be performed for biomass (cereals and oil crops), diversity (oil crops) and enzymatic activity (oil crops)

```{r, echo = FALSE}

# Biomass

kable(table(biomass$magpie_class, biomass$agricultural_system), caption = "Metric: biomass")  %>%
  kable_styling( full_width = F, position = "float_left") 

# Development

kable(table(development$magpie_class, development$agricultural_system), caption = "Metric: development") %>%
  kable_styling( full_width = F, position = "right")

# Diversity

kable(table(diversity$magpie_class, diversity$agricultural_system), caption = "Metric: diversity") 

# Efficiency
  
kable(table(efficiency$magpie_class, efficiency$agricultural_system), caption = "Metric: efficiency") %>%
  kable_styling( full_width = F, position = "float_left")

# Enzymatic activity

kable(table(Enzymatic_activity$magpie_class, Enzymatic_activity$agricultural_system), caption = "Metric: Enzymatic_activity") %>%
  kable_styling( full_width = F, position = "right")

# Reproduction

kable(table(reproduction$magpie_class, reproduction$agricultural_system), caption = "Metric: reproduction") %>%
  kable_styling( full_width = F, position = "float_left")

# Survival

kable(table(survival$magpie_class, survival$agricultural_system), caption = "Metric: survival") %>%
  kable_styling( full_width = F, position = "right")

```

### 3.5 Biomass, diversity and enzymatic activity analysis per **magpie class**: cereals and oil crops

An analysis could be performed for biomass (cereals and oil crops), diversity (oil crops) and enzymatic activity (oil crops)

```{r}
# Biomass: Cereals 

cereals <- subset(d, magpie_class == "Cereals" & biodiveristy_metric_category == "biomass")

unique(cereals$Paper_ID)

```
There are four papers that provide data for biomass for the cereals magpie class

```{r}

biomass_cereals <- blmer(LRR ~  Treatment + (1|ID), weights = Weight, data = cereals, control = lmerControl(optimizer ="Nelder_Mead"))

summary(biomass_cereals)

```

```{r}
# Biomass: Oil crops

oil_crops <- subset(d, magpie_class == "Oil crops" & biodiveristy_metric_category == "biomass")

unique(oil_crops$Paper_ID)
```

There are two papers that provide data for biomass for the oil crops magpie class

```{r}

biomass_oil_crops <- blmer(LRR ~  Treatment + (1|ID), weights = Weight, data = oil_crops, control = lmerControl(optimizer ="Nelder_Mead"))

summary(biomass_oil_crops)

```

```{r}
# Diversity: Oil crops

oil_crops <- subset(d, magpie_class == "Oil crops" & biodiveristy_metric_category == "diversity")

unique(oil_crops$Paper_ID)
```

There are nine papers that provide data for diversity for the oil crops magpie class

```{r}

diversity_oil_crops <- blmer(LRR ~  Treatment + (1|ID), weights = Weight, data = oil_crops, control = lmerControl(optimizer ="Nelder_Mead"))

summary(diversity_oil_crops)

```
```{r}
# Enzymatic activity: Oil crops

oil_crops <- subset(d, magpie_class == "Oil crops" & biodiveristy_metric_category == "Enzymatic_activity")

unique(oil_crops$Paper_ID)

```

There are two papers that provide enzymatic activity data for the oil crops magpie class

```{r}

enzymatic_oil_crops <- blmer(LRR ~  Treatment + (1|ID), weights = Weight, data = oil_crops, control = lmerControl(optimizer ="Nelder_Mead"))

summary(enzymatic_oil_crops)

```

